---
title: "p8105_hw2_yo2336"
author: "Yoo Rim Oh"
output: github_document
---
```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

## Read Mr.Trash excel data and clean

Named the data frame as trash_df and set the range to omit the note sections without any data.
Removed all missing values (non-dumpster specific data) under dumpster column and mutated
sports_balls column to integer from doubles.

```{r}
trash_df = 
  read_excel("data/trash_wheel_updated.xlsx", sheet = "Mr. Trash Wheel", 
             range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(sports_balls))
```


## Read Precipitation Data (2018 & 2019) then clean and bind

Named the data frame as precip_df_2018 and precip_df_2019 and set the range to only 
show observations and not totals. Renamed the Total column to precipitation_in 
and added an additional column for year data was collected for both data sets.
Arranged both data sets so that year variable comes first.

Combined two data sets using bind_rows and changed month column to convert numeric 
form of month to character month names.

```{r}
precip_df_2018 =
  read_excel("data/trash_wheel_updated.xlsx", sheet = "2018 Precipitation", 
             range = "A2:B14") %>%
  janitor::clean_names() %>%
  rename(precipitation_in = total) %>%
  mutate(year = 2018) %>%
  relocate(year)


precip_df_2019 =
  read_excel("data/trash_wheel_updated.xlsx", sheet = "2019 Precipitation", 
             range = "A2:B14") %>%
  janitor::clean_names() %>%
  rename(precipitation_in = total) %>%
  mutate(year = 2019) %>%
  relocate(year)


precip_df =
  bind_rows(precip_df_2018, precip_df_2019) %>%
  janitor::clean_names() %>%
  mutate(month = month.name[as.numeric(month)])
```


## Trash Wheel and Precipitation Data Description
```{r}
ncol(trash_df)
names(trash_df)
nrow(trash_df)


ncol(precip_df)
names(precip_df)
nrow(precip_df)
nrow(precip_df_2018)
nrow(precip_df_2019)


sum(pull(precip_df_2018, precipitation_in))

median(pull(filter(trash_df, year == 2019), sports_balls))
```

The Mr.Trash Wheel Data has 14 columns representing variables such as dumpster number, 
month/year/date that specific dumpster was collected, weight and volume of the dumpster, 
8 different types of trash categories, and homes powered. It has total of 454 rows 
representing the total number of observations, or dumpster specific data, collected 
between May 2014 to January 2021.

The combined Precipitation Data during 2018 and 2019 has 3 columns for year and month of when 
the data was collected and precipitation in inches. It has total of 24 rows representing the total
number of observations taken every month each year.

The total precipitation in 2018 was 70.33 in.
The median number of sports balls found in a dumpster in 2019 was 9.


# Problem 2


## Part 1 Read and clean pols-month.csv

Named the data frame as pols_month_df. Reorganized the date to have individual columns
for year, month, and day each and converted month to character month name.
Reorganized so prez_gop and prez_dem lined up together in order to use pivot_longer.
Changed column names to gop and dem instead of prez_gop and prez_dem so when combined 
together, any observation not 0 as its value will be reassigned to new values either 
gop or dem. Rearranged so data set begins with year, month, and president variables.

```{r}
pols_month_df =
  read_csv("data/fivethirtyeight_datasets/pols_month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = month.name[as.numeric(month)]) %>%
  relocate(year, month, prez_gop, prez_dem) %>%
  rename(gop = prez_gop, dem = prez_dem) %>%
  pivot_longer(
    gop:dem,
    names_to = "president",
    values_to = "prez_numeric") %>%
  filter(prez_numeric != 0) %>%
  select(-c("prez_numeric", "day")) %>%
  relocate(year, month, president)
```

## Part 2 Read and clean snp.csv

Named data frame as snp_df. Used lubridate::mdy() to convert date in m/d/yy form to
yyyy-mm-dd format then separated to create columns for each year, month, and day.
Converted month variable to full month name character, removed day column, and rearranged 
according to year and month.

```{r}
snp_df =
  read_csv("data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  mutate(date = lubridate::mdy(date)) %>%
  separate(date, into = c("year", "month", "day")) %>%
  arrange(year, month) %>%
  mutate(month = month.name[as.numeric(month)]) %>%
  select(-"day")
```

## Part 3 Read and clean unemployment.csv

Named the data frame as unemp_df and used pivot_longer to transform individual month 
columns to one column named month. Converted abbreviated month names to numeric then to 
full month name character.

```{r}
unemp_df =
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "percent_unemployment") %>%
  mutate(month = month.name[match(month, month.abb)]) %>%
  janitor::clean_names()
```

## Join snp and pols data

Combined pols_month and snp using left join by year and month variables.
Converted year column as numeric (doubles column) instead of character column.

```{r}
snp_pols_df =
  left_join(pols_month_df, snp_df, by = c("year", "month")) %>%
  janitor::clean_names() %>%
  mutate(year = as.numeric(year))
```

## Join combined snp_pols with unemployment

Combined the above snp_pols joined data to unemp data using left_join by year and month.

```{r}
snp_pols_unemp_df =
  left_join(snp_pols_df, unemp_df, by = c("year", "month")) %>%
  janitor::clean_names()
```

## FiveThirtyEight data description
```{r}
ncol(pols_month_df)
names(pols_month_df)
nrow(pols_month_df)
range(pull(pols_month_df, year))

ncol(snp_df)
names(snp_df)
nrow(snp_df)
range(pull(snp_df, year))

ncol(unemp_df)
names(unemp_df)
nrow(unemp_df)
range(pull(snp_df, year))
```

The pols_month data has 9 columns representing variables such as year/month, 
president for which party (democrat as dem and republican as gop) that year's 
president belonged, and number of each party's governors, senators, and representatives 
at the time. It has 822 rows for total number of observations ranging from the years 
1947 to 2015 every month.

The snp data has 3 columns representing variables year, month, and close for 
closing values of the S&P stock index at the time. It has 787 rows for total number 
of observations ranging from the years 1969 to 2068 every month.

The unemployment data has 3 columns representing variables year, month, and percent 
unemployment at the time of data collection. It has 816 rows for total number 
of observations ranging from the years 1969 to 2068 every month.


# Problem 3

## Load and clean baby_name data

Named data frame as baby_name_df and used unique() to remove any duplicate observations.
Rearranged so data started with year, rank, and name and arranged according to year and ethnicity.
Recoded so all ethnicity variables are unified and changed ethnicity and childs_first_name
values to be in lower case.

```{r}
baby_name_df =
  read_csv("data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  unique() %>%
  relocate(year_of_birth, rank, childs_first_name) %>%
  arrange(year_of_birth, ethnicity) %>%
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", 
                  "BLACK NON HISP" = "BLACK NON HISPANIC", 
                  "WHITE NON HISP" = "WHITE NON HISPANIC"),
         ethnicity = str_to_lower(ethnicity),
         childs_first_name = str_to_lower(childs_first_name))
```

## Create table for name Olivia

Created a new data frame named baby_name_tbl for name Olivia only. Used pivot_wider to 
assign individual columns for each ethnicity and the ranks to be their values.
Removed all irrelevant columns for name Olivia.
Created data frame for each ethnicity separately and removed all missing values. Then
joined the 4 ethnicity based data frames together back again to show ranks of name 
Olivia for each ethnicity each year.
Created table using knitr::kable of the joined data frame.

```{r}
baby_name_tbl = 
    pivot_wider(baby_name_df,
           names_from = "ethnicity",
           values_from = "rank") %>%
    filter(childs_first_name == "olivia") %>%
    select(-c("childs_first_name", "gender", "count"))

aapi_olivia = select(baby_name_tbl, year_of_birth, "asian and pacific islander") %>%
  drop_na("asian and pacific islander")
bnh_olivia = select(baby_name_tbl, year_of_birth, "black non hispanic") %>%
  drop_na("black non hispanic")
h_olivia = select(baby_name_tbl, year_of_birth, "hispanic") %>%
  drop_na("hispanic")
wnh_olivia = select(baby_name_tbl, year_of_birth, "white non hispanic") %>%
  drop_na("white non hispanic")

olivia_tbl = left_join(aapi_olivia, left_join(bnh_olivia, left_join(h_olivia, wnh_olivia)))
knitr::kable(olivia_tbl, caption = "Rank in Popularity for Baby's Name Olivia")
```

## Create table for popular male name Ethan

Same flow as table for Olivia above except for baby name Ethan.

```{r}
male_name_tbl = 
    pivot_wider(baby_name_df,
           names_from = "ethnicity",
           values_from = "rank") %>%
    filter(childs_first_name == "ethan") %>%
    select(-c("childs_first_name", "gender", "count"))

aapi_ethan = select(male_name_tbl, year_of_birth, "asian and pacific islander") %>%
  drop_na("asian and pacific islander")
bnh_ethan = select(male_name_tbl, year_of_birth, "black non hispanic") %>%
  drop_na("black non hispanic")
h_ethan = select(male_name_tbl, year_of_birth, "hispanic") %>%
  drop_na("hispanic")
wnh_ethan = select(male_name_tbl, year_of_birth, "white non hispanic") %>%
  drop_na("white non hispanic")

ethan_tbl = left_join(aapi_ethan, left_join(bnh_ethan, left_join(h_ethan, wnh_ethan)))
knitr::kable(ethan_tbl, caption = "Rank in Popularity for Baby's Name Ethan")
```

## Scatter plot for names of white non hispanic male born in 2016

Created a data frame named white_male_2016 first with only male names from 2016 ranked for 
white non hispanic group. Then used ggplot()/geom_point to create scatter plot with rank 
on x-axis and count on y-axis.

```{r}
white_male_2016 =
  filter(baby_name_df, gender == "MALE", 
         ethnicity == "white non hispanic", 
         year_of_birth == 2016)

ggplot(white_male_2016, aes(x = rank, y = count)) +
  geom_point() +
    labs(title = "Popular Names of White Non-Hispanic Male Born in 2016")
```

